---
title: "Joint Random Variables: Worked Example"
subtitle: "Discrete Probability Distributions"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(tidyquant)
library(ggside)
library(ggplot2)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#0cB2A3",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE,eval=FALSE,include=FALSE}

## ASI - CT3 - 2005 - June - Question 08

```

To test if the observed frequencies of birds in six categories are compatible with the assumption that the park is visited by the same proportion of birds in the population, we can use the chi-squared (χ²) test.

Here are the steps involved:

1. **State the null hypothesis (H₀)**:
   - H₀: The park is visited by birds in the same proportion in each category.

2. **Calculate the expected frequencies (E)**:
   - Since we assume the park is visited by birds in the same proportion, we can divide the total number of birds by the number of categories.
   - Total number of birds = 6 + 7 + 13 + 14 + 9 + 5 = 54
   - Number of categories = 6
   - Expected frequency for each category (E) = Total number of birds / Number of categories = 54 / 6 = 9

3. **Calculate the chi-squared (χ²) statistic**:
   - χ² = Σ ((Observed frequency - Expected frequency)² / Expected frequency)
   - χ² = ((6-9)²/9) + ((7-9)²/9) + ((13-9)²/9) + ((14-9)²/9) + ((9-9)²/9) + ((5-9)²/9)
   - χ² = (9/9) + (4/9) + (16/9) + (25/9) + (0/9) + (16/9)
   - χ² = 1 + 0.444 + 1.778 + 2.778 + 0 + 1.778
   - χ² = 7.778

4. **Determine the degrees of freedom (df)**:
   - Degrees of freedom (df) = Number of categories - 1 = 6 - 1 = 5

5. **Find the critical value from the chi-squared distribution table**:
   - At 5% level of significance (α = 0.05) and 5 degrees of freedom (df = 5), the critical value (χ²₀.₀₅) is approximately 11.07.

6. **Compare the calculated χ² value with the critical value**:
   - Calculated χ² = 7.778
   - Critical χ²₀.₀₅ = 11.07

7. **Conclusion**:
   - Since the calculated χ² value (7.778) is less than the critical value (11.07), we fail to reject the null hypothesis (H₀).

Therefore, the observed frequencies of birds in the six categories are compatible with the assumption that the park is visited by the same proportion of birds belonging to these categories in the population.

If you have any more questions or need further assistance, just let me know! 🦜