---
title: "Joint Random Variables: Worked Example"
subtitle: "Discrete Probability Distributions"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(tidyquant)
library(ggside)
library(ggplot2)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#0cB2A3",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE,eval=FALSE,include=FALSE}

## ASI - CT3 - 2005 - June - Question 05

```

### Part (a)

**Central Limit Theorem (CLT) for Independent Identically Distributed Random Variables:**

The Central Limit Theorem states that if $X_1, X_2, ..., X_n$ are independent and identically distributed (i.i.d.) random variables with mean $\mu$ and variance $\sigma^2$, then the distribution of the sample mean $\bar{X} = \frac{1}{n} \sum_{i=1}^{n} X_i$ approaches a normal distribution as the sample size $n$ becomes large. Specifically,
$$ \bar{X} \approx N\left(\mu, \frac{\sigma^2}{n}\right) $$
---

**Given:**
- Mean, $\mu = 167$
- Standard deviation, $\sigma = 27$
- Sample size, $n = 36$

**Find:**
- Probability that the sample mean lies between 163 and 171 units.

First, compute the standard error of the mean:
$$ \text{SE} = \frac{\sigma}{\sqrt{n}} = \frac{27}{\sqrt{36}} = \frac{27}{6} = 4.5 $$

---

The sample mean $\bar{X}$ follows the normal distribution:
$$ \bar{X} \sim N\left(167, 4.5^2\right) $$

To find the probability:
$$ P(163 < \bar{X} < 171) = P\left(\frac{163 - 167}{4.5} < Z < \frac{171 - 167}{4.5}\right) $$
$$ = P\left(\frac{-4}{4.5} < Z < \frac{4}{4.5}\right) $$
$$ = P(-0.8889 < Z < 0.8889) $$
---

From the standard normal distribution table:
$$ P(Z < 0.8889) = 0.8133 $$
$$ P(Z < -0.8889) = 1 - 0.8133 = 0.1867 $$

Therefore:
$$ P(-0.8889 < Z < 0.8889) = 0.8133 - 0.1867 = 0.6266 $$

Thus, the approximate probability that the sample mean lies between 163 and 171 units is $0.6266$.

---

### Part (b)

**Given:**
- The distribution function $F_X(x)$ of $X$ is:
$$ F_X(x) = 
   \begin{cases} 
   0 & \text{if } x < 0 \\
   3x^2 - 2x^3 & \text{if } 0 \le x \le 1 \\
   1 & \text{if } x > 1 
   \end{cases}
$$

---

The volume $V$ of the sphere of radius $X$ is:
$$ V = \frac{4}{3}\pi X^3 $$

**Find the probability that the volume of the sphere is greater than $V_0$.**

$$ P(V > V_0) = P\left(\frac{4}{3}\pi X^3 > V_0\right) $$

$$ P\left(X^3 > \frac{3V_0}{4\pi}\right) $$

---

$$ P\left(X > \left(\frac{3V_0}{4\pi}\right)^{1/3}\right) $$

Let $y = \left(\frac{3V_0}{4\pi}\right)^{1/3}$.

---
**For what values of $V_0$ is the above probability zero?**

The probability is zero if $y > 1$, as $X$ lies between 0 and 1.

$$\left(\frac{3V_0}{4\pi}\right)^{1/3} > 1$$

Cubing both sides:
$$\frac{3V_0}{4\pi} > 1$$
---
Solving for $V_0$:
$$ V_0 > \frac{4\pi}{3} $$

---

### Conclusion

- **Part (a):** 

The approximate probability that the sample mean lies between 163 and 171 units is $0.6266$.

- **Part (b):**

  - $P(V > V_0) = 1 - \left( \frac{3V_0}{4\pi} \right) + 2 \left( \frac{3V_0}{4\pi} \right)^2$ for $0 \le V_0 \le \frac{4\pi}{3}$.
  
  - The probability that the volume of the sphere is greater than $V_0$ is zero if $V_0 > \frac{4\pi}{3}$.

---