---
title: "Joint Random Variables: Worked Example"
subtitle: "Discrete Probability Distributions"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(tidyquant)
library(ggside)
library(ggplot2)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#0cB2A3",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE,eval=FALSE,include=FALSE}

## ASI - CT3 - 2005 - November - Question 09

```


The joint pmf of $X$ and $Y$ is given below.

| Y | X = 0 | X = 1 | X = 2 |
|---|-------|-------|-------|
| 0 | 1/6   | 1/3   | 1/12  |
| 1 | 2/9   | 1/6   | 0     |
| 2 | 1/36  | 0     | 0     |

Find
1. **Mean values of $X$ and $Y$** 
2. **Find the covariance between $X$ and $Y$** 
3. **Conditional distribution of X given Y = 1** 
4. **Correlation between $X$ and $Y$**

---


### a) Mean values of $X$ and $Y$

We need to compute $E(X)$ and $E(Y)$ using the given joint pmf:

$$
\begin{array}{c|ccc}
  & Y = 0 & Y = 1 & Y = 2 \\
\hline
X = 0 & \frac{1}{6} & \frac{1}{3} & \frac{1}{12} \\
X = 1 & \frac{2}{9} & \frac{1}{6} & 0 \\
X = 2 & \frac{1}{36} & 0 & 0 \\
\end{array}
$$

1. **Mean of $X$**:
$$
E(X) = \sum_x \sum_y x \cdot P(X=x, Y=y)
$$
$$
E(X) = 0 \cdot \left( \frac{1}{6} + \frac{1}{3} + \frac{1}{12} \right) + 1 \cdot \left( \frac{2}{9} + \frac{1}{6} \right) + 2 \cdot \frac{1}{36}
$$
$$
E(X) = 1 \cdot \left( \frac{2}{9} + \frac{1}{6} \right) + 2 \cdot \frac{1}{36}
$$
$$
E(X) = \frac{2}{9} + \frac{1}{6} + \frac{2}{36}
$$
$$
E(X) = \frac{2}{9} + \frac{1}{6} + \frac{1}{18}
$$
$$
E(X) = \frac{4}{18} + \frac{3}{18} + \frac{1}{18}
$$
$$
E(X) = \frac{8}{18} = \frac{4}{9}
$$

2. **Mean of $Y$**:
$$
E(Y) = \sum_x \sum_y y \cdot P(X=x, Y=y)
$$
$$
E(Y) = 0 \cdot \left( \frac{1}{6} + \frac{2}{9} + \frac{1}{36} \right) + 1 \cdot \left( \frac{1}{3} + \frac{1}{6} \right) + 2 \cdot \frac{1}{12}
$$
$$
E(Y) = 1 \cdot \left( \frac{1}{3} + \frac{1}{6} \right) + 2 \cdot \frac{1}{12}
$$
$$
E(Y) = \frac{1}{3} + \frac{1}{6} + \frac{2}{12}
$$
$$
E(Y) = \frac{2}{6} + \frac{1}{6} + \frac{1}{6}
$$
$$
E(Y) = \frac{4}{6} = \frac{2}{3}
$$

### b) Covariance between $X$ and $Y$

Covariance is given by:
$$
\text{Cov}(X, Y) = E(XY) - E(X)E(Y)
$$

First, we need to compute $E(XY)$:

$$
E(XY) = \sum_x \sum_y xy \cdot P(X=x, Y=y)
$$
$$
E(XY) = 0 \cdot \left( 0 + 0 + 0 \right) + 1 \cdot \left( 0 + \frac{1}{6} \right) + 2 \cdot \left( \frac{1}{6} \cdot 1 + \frac{1}{36} \cdot 2 \right)
$$
$$
E(XY) = 0 + \frac{1}{6} + \left( \frac{1}{6} \cdot 2 + \frac{1}{36} \cdot 4 \right)
$$
$$
E(XY) = \frac{1}{6} + \frac{2}{12} + \frac{4}{36}
$$
$$
E(XY) = \frac{1}{6} + \frac{1}{6} + \frac{1}{9}
$$
$$
E(XY) = \frac{2}{6} + \frac{1}{9}
$$
$$
E(XY) = \frac{3}{9} + \frac{1}{9}
$$
$$
E(XY) = \frac{4}{9}
$$

Now, compute the covariance:
$$
\text{Cov}(X, Y) = \frac{4}{9} - \left( \frac{4}{9} \cdot \frac{2}{3} \right)
$$
$$
\text{Cov}(X, Y) = \frac{4}{9} - \frac{8}{27}
$$
$$
\text{Cov}(X, Y) = \frac{12}{27} - \frac{8}{27}
$$
$$
\text{Cov}(X, Y) = \frac{4}{27}
$$

### c) Conditional Distribution of $X$ given $Y = 1$

We need to find $P(X=x | Y=1)$:

$$
P(X=x | Y=1) = \frac{P(X=x, Y=1)}{P(Y=1)}
$$

First, find $P(Y=1)$:
$$
P(Y=1) = \sum_x P(X=x, Y=1) = \frac{1}{3} + \frac{1}{6} + 0 = \frac{1}{2}
$$

Now, calculate the conditional probabilities:
$$
P(X=0 | Y=1) = \frac{P(X=0, Y=1)}{P(Y=1)} = \frac{1/3}{1/2} = \frac{2}{3}
$$
$$
P(X=1 | Y=1) = \frac{P(X=1, Y=1)}{P(Y=1)} = \frac{1/6}{1/2} = \frac{1}{3}
$$
$$
P(X=2 | Y=1) = \frac{P(X=2, Y=1)}{P(Y=1)} = 0
$$

### d) Correlation between $X$ and $Y$

The correlation $\rho$ is given by:
$$
\rho(X, Y) = \frac{\text{Cov}(X, Y)}{\sigma_X \sigma_Y}
$$

We need the standard deviations $\sigma_X$ and $\sigma_Y$:
$$
\sigma_X = \sqrt{E(X^2) - (E(X))^2}
$$
$$
E(X^2) = \sum_x \sum_y x^2 \cdot P(X=x, Y=y) = 0^2 \cdot ( \frac{1}{6} + \frac{1}{3} + \frac{1}{12} ) + 1^2 \cdot ( \frac{2}{9} + \frac{1}{6} ) + 2^2 \cdot \frac{1}{36}
$$
$$
E(X^2) = 0 + \left( \frac{2}{9} + \frac{1}{6} \right) + 4 \cdot \frac{1}{36}
$$
$$
E(X^2) = \frac{2}{9} + \frac{3}{18} + \frac{4}{36}
$$
$$
E(X^2) = \frac{4}{18} + \frac{3}{18} + \frac{2}{18}
$$
$$
E(X^2) = \frac{9}{18}
$$
$$
E(X^2) = \frac{1}{2}
$$

$$
\sigma_X = \sqrt{\frac{1}{2} - \left( \frac{4}{9} \right)^2} = \sqrt{\frac{1}{2} - \frac{16}{81}} = \sqrt{\frac{40.5}{81} - \frac{16}{81}} = \sqrt{\frac{24.5}{81}} = \frac{\sqrt{24.5}}{9}
$$

Similarly, for $\sigma_Y$:
$$
\sigma_Y = \sqrt{E(Y^2) - (E(Y))^2}
$$
$$
E(Y^2) = \sum_x \sum_y y^2 \cdot P(X=x, Y=y) = 0^2 \cdot ( \frac{1}{6} + \frac{2}{9} + \frac{1}{36} ) + 1^2 \cdot ( \frac{1}{3} + \frac{1}{6} ) + 2^2 \cdot \frac{1}{12}
$$
$$
E(Y^2) = 0 + \left( \frac{1}{3} + \frac{1}{6} \right)