---
title: "Joint Random Variables: Worked Example"
subtitle: "Discrete Probability Distributions"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(tidyquant)
library(ggside)
library(ggplot2)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#0cB2A3",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r echo=FALSE,eval=FALSE,include=FALSE}

## ASI - CT3 - 2005 - November - Question 07

```


A random variable ${\displaystyle X}$ follows the hypergeometric distribution if its probability mass function is given by

$$ {\displaystyle \Pr(X=k)={\frac {{\binom {K}{k}}{\binom {N-K}{n-k}}}{\binom {N}{n}}},}$$

\noindent where

\begin{itemize}
    \item     ${\displaystyle N} $ is the population size,
    \item       ${\displaystyle K}$  is the number of success states in the population,
    \item       ${\displaystyle n}$  is the number of draws (i.e. quantity drawn in each trial),
     \item      ${\displaystyle k}$ is the number of observed successes,
     \item     ${\textstyle \textstyle {a \choose b}}$ is a binomial coefficient.

\end{itemize}


\end{framed}
\medskip
\noindent \textbf{Solution}\\
Invoking Hyper geometric distribution with
x = 2, n = 5, N = 120 and K = 80,


$$ {\displaystyle \Pr(X=2)={\frac {{\binom {80}{2}}{\binom {40}{3}}}{\binom {120}{5}}},}$$


\begin{itemize}
    \item ${ \displaystyle {\binom {80}{2}}   =  \frac{80 \times 79}{2\times 1}  = (40 \times 79) }$
    \item ${ \displaystyle  {\binom {40}{2}}  =  \frac{40 \times 39 \times 38}{3 \times 2\times 1}  = (20 \times 13 \times 38) }$
    \item ${ \displaystyle {\binom {120}{5}}  =  \frac{120 \times 119 \times 118 \times 117 \times 116 }{5 \times 4 \times 3 \times 2\times 1} = (119 \times 118 \times 117 \times 116)}$
\end{itemize}

\begin{eqnarray*}
\Pr(X=2) &=& {\frac {{\binom {80}{2}}{\binom {40}{3}}}{\binom {120}{5}}} \\
& &  \\
&=& \frac{(40 \times 79) \times (20 \times 13 \times 38) }{(119 \times 118 \times 117 \times 116)}\\
&=&  \left( \frac{40 \times 79}{119 \times 118} \right)  \times  \left( \frac{20 \times 13 \times 38}{117 \times 116} \right) \\
& & \\
&=& \frac{3,160}{14,042}  \times \frac{9,880}{13,572}\\
& & \\
&=& 0.22503 \times 0.73533\\
& & \\
&=& 0.16547 \\
\end{eqnarray*}
