---
title: "Joint Random Variables: Worked Example"
subtitle: "Discrete Probability Distributions"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(tidyquant)
library(ggside)
library(ggplot2)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#0cB2A3",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE,eval=FALSE,include=FALSE}

## ASI - CT3 - 2005 - November - Question 02

```

1. **Given data**:
   - Mean of 5 observations (x̄) = 15
   - Variance of 5 observations (s²) = 9
   - Number of observations (n) = 5

2. **Calculate the sum of the original observations (Σx)**:
   - Σx = n * x̄
   - Σx = 5 * 15
   - Σx = 75

3. **Add the two new observations (-3 and 10)**:
   - Σx (new) = 75 - 3 + 10
   - Σx (new) = 82

4. **Calculate the new mean (x̄₇)**:
   - Number of new observations = 7
   - New mean (x̄₇) = Σx (new) / 7
   - New mean (x̄₇) = 82 / 7
   - New mean (x̄₇) ≈ 11.71

5. **Calculate the corrected sum of squares (Σx²)**:
   - Variance formula: s² = (Σ(xi - x̄)²) / (n - 1)
   - Rearrange the formula: Σ(xi - x̄)² = s² * (n - 1)
   - Σ(xi - x̄)² = 9 * (5 - 1)
   - Σ(xi - x̄)² = 9 * 4
   - Σ(xi - x̄)² = 36

6. **Calculate the sum of squares of new observations (Σx² new)**:
   - Σx² new = Σ(xi - x̄)² + (-3)² + 10²
   - Σx² new = 36 + 9 + 100
   - Σx² new = 145

7. **Calculate the new variance (s²₇)**:
   - New variance (s²₇) = (Σx² new - (Σx new)² / n) / (n - 1)
   - New variance (s²₇) = (145 - (82)² / 7) / 6
   - New variance (s²₇) ≈ (145 - 6724 / 7) / 6
   - New variance (s²₇) ≈ (145 - 960.57) / 6
   - New variance (s²₇) ≈ (-815.57) / 6
   - New variance (s²₇) ≈ 51.5 (approximately)

So, the new mean of the 7 observations is approximately **11.71** and the new variance is approximately **51.5**.

I hope this helps! If you have any other questions or need further clarification, just let me know. 😊