---
title: "Joint Random Variables: Worked Example"
subtitle: "Discrete Probability Distributions"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(tidyquant)
library(ggside)
library(ggplot2)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#0cB2A3",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



The number of claims, which arise in a year under each individual policy of a certain
type, has a Poisson distribution with unknown mean l. Last year’s experience for a
group of independent policies of this type is examined and it is found that no claims
arose under 78% of the policies (and claims did arise under 22% of the policies).
Find the maximum likelihood estimate of l

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Solution
\medskip
\noident \textbf{Solution}\\
P(no claims) = e -l 
Observed proportion of “no claims” = 0.78 
\ MLE of el is 0.78 so MLE of l is –log(0.78) = 0.248

---

Let's find the maximum likelihood estimate (MLE) of $\lambda$, the mean of the Poisson distribution.

Given that the number of claims follows a Poisson distribution with mean $\lambda$, the probability of observing zero claims under a policy is:

$$ P(\text{no claims}) = e^{-\lambda} $$

From the given information, the observed proportion of policies with no claims is 0.78. This observed proportion serves as the MLE for $e^{-\lambda}$.

Hence, we have:

$$ e^{-\lambda} = 0.78 $$

Taking the natural logarithm on both sides to solve for $\lambda$:

$$ -\lambda = \log(0.78) $$
$$ \lambda = -\log(0.78) $$

Computing this value:

$$ \lambda \approx -(-0.248) $$
$$ \lambda \approx 0.248 $$

Therefore, the maximum likelihood estimate (MLE) of $\lambda$ is approximately $0.248$.