---
title: "Joint Random Variables: Worked Example"
subtitle: "Discrete Probability Distributions"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(tidyquant)
library(ggside)
library(ggplot2)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#0cB2A3",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Sure, let's walk through the process of calculating the lower 95% confidence interval for λ using the provided data.

### Step-by-Step Solution

1. **Data and Assumptions:**
   - Sample size, $n = 200$
   - Total claims, $X = 52$
   - Poisson distribution is assumed for the number of claims.

2. **Estimate the Mean:**
   - The mean number of claims per policy, $\hat{\lambda}$, is estimated by the sample mean:
     $$
     \hat{\lambda} = \frac{X}{n} = \frac{52}{200} = 0.26
     $$

3. **Using Normal Approximation:**
   - The distribution of the sample mean $\hat{\lambda}$ can be approximated by a normal distribution with mean $\lambda$ and variance $\frac{\lambda}{n}$.
     $$
     \hat{\lambda} \sim N \left( \lambda, \frac{\lambda}{n} \right)
     $$

4. **Standard Error:**
   - The standard error (SE) of $\hat{\lambda}$ is:
     $$
     \text{SE} = \sqrt{\frac{\hat{\lambda}}{n}} = \sqrt{\frac{0.26}{200}} \approx 0.036
     $$

5. **Z-value for 95% Confidence:**
   - The Z-value for a lower 95% confidence interval (one-tailed) is 1.645.

6. **Lower Confidence Limit:**
   - The lower limit of the 95% confidence interval is given by:
     $$
     \hat{\lambda} - Z \times \text{SE} = 0.26 - 1.645 \times 0.036 \approx 0.26 - 0.059 = 0.201
     $$

### Conclusion:
The lower limit of the 95% confidence interval for $\lambda$ is approximately $0.201$.

Therefore, the lower 95% confidence interval for $\lambda$ is $(0.201, \infty)$.

If you need further clarification or have any other questions, feel free to ask!