\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
\begin{enumerate}
\item A random sample of fifty claim amounts (£) arising in a particular section of an
insurance company's business are displayed below in a stem and leaf plot:

\begin{verbatim}
   15
16
17
18
19
20
21
22
23
24
25
26
27
28
29
30
31
14678
0233368889
0000001233457888
3456779
0257
03
07
3
38
2
Stem unit = 100
Leaf unit = 10 
\end{verbatim}

The sum of the fifty amounts (before rounding) is £92780.
Calculate the mean and median claim amounts. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent \textbf{solution}
1 Mean = 92780/50 = £1855.60
25.5th value in order = (17.3 + 17.4)/2 = 17.35 so median amount = £1735
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\item Consider a random sample of 47 white-collar workers and a random sample of 24 blue-collar workers from the workforce of a large company. \begin{itemize}
    \item The mean salary for
the sample of white-collar workers is £28,470 and the standard deviation is £4,270; \item whereas the mean salary for the sample of blue-collar workers is £21,420
and the standard deviation is £3,020.
\end{itemize}
Calculate the mean and the standard deviation of the salaries in the combined
sample of 71 employees. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
2 For simplicity change the units into thousands. 
\begin{itemize}
    \item The sum of the data is
\[\sum x = 47 \times 28.47 + 24 \times  21.42 = 1,852.17.\]
\item So, the mean salary is 1852.17
71 = 26.087, or £26,087.
\item The sum of squares is given by
\[ \sum x^2 = \{46 \times 4.272 + 47 \times 28.472\} + \{23 \times 3.022 + 24 \times 21.422\} = 50,155.5.\]
\item Thus, the standard deviation of the 71 values is
50155.5 1852.172 71
70
−
= 5.124, or £5,124.
\end{itemize}



\end{document}
