\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
\begin{enumerate}%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
16 At the end of the skiing season the tourist board in a mountain region examines
the records of ten ski resorts. For each one it obtains the total number (y,
thousands) of visitor-days during the season as a measure of the resort’s
popularity, and the ski-lift capacity (x, thousands), being the maximum number
of skiers that can be transported per hour. The resulting data are given in the
following table:
\begin{verbatim}
    Resort: A B C D E F G H I J
Lift capacity x: 1.9 3.3 1.2 4.2 1.5 2.2 1.0 5.6 1.9 3.8
Visitor-days y: 15.1 22.6 9.2 37.5 8.9 21.1 5.8 41.0 9.2 32.4
\sum x = 26.6, \sum x2 = 91.08, \sum y = 202.8, \sum y2 = 5603.12, \sum xy = 707.58
\end{verbatim}

\begin{enumerate}[(a)]
\item Draw a scatterplot of y against x and comment briefly on any relationship
between a resort’s popularity and its ski-lift capacity. 
\item Calculate the correlation coefficient between x and y and comment briefly
in the light of your comment in part (i). 
\item Calculate the fitted linear regression equation of y on x. 
\item (a) Calculate the “total sum of squares” together with its partition into
the “regression sum of squares” and the “residual sum of squares”.
(b) Use the values in part (iv)(a) above to calculate the coefficient of
determination R2 and comment briefly on its relationship with the
correlation coefficient calculated in part (ii).
(c) Use the values in part (iv)(a) above to calculate an estimate of the
error variance $\sigma^2$ in the usual linear regression model. 
\item Suppose that a resort can increase its ski-lift capacity by 500 skiers per
hour.
\end{enumerate}
Estimate the increase in the number of visitor-days it can expect in a
season, and specify a standard error for this estimate.
%%%%%%% Question 14

% ii 

16 (i)
Seems to be a strong increasing linear relationship.
\begin{itemize}
\item
${ \displaystyle Sxx = 91.08 − \frac{26.6^2}{10} = 20.324  }$
\item ${ \displaystyle Syy = 5603.12 − \frac{202.8^2}{10} = 1490.336  }$
\item ${ \displaystyle Sxy = 707.58 −  − \frac{(26.6)(202.8)}{10} = 168.132  }$
\item ${ \displaystyle r = \frac{168.132}{\sqrt{(20.324)(1490.336)}}= 0.966}$
\end{itemize}
Large and positive agreeing with comment above.
Large and positive agreeing with comment above.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Page 10
\begin{itemize}
    \item 
(iii) y = $\alpha$ˆ + βˆ x where
ˆβ
=
168.132
20.324
= 8.2726
−1.73 + 8.273x
ˆ$\alpha$ =
202.8 26.6
8.2726
10 10
  −  
 
= −1.73
\item (iv) (a) SSTOT = Syy = 1490.336
SSRES = Syy −
2
xy
xx
S
S
(formula from green book)
= 1490.336 −
(168.132)2
20.324
= 99.450
∴ SSREG = 1390.886\\
(b) R2 =
SSREG
SSTOT
= 0.933 = r2 from (ii)\\
(c) \sum ˆ 2 =
1
8
(SSRES) = 12.43
\item (v) If x increases by 0.5 (500 in ‘000’s), then expected increase in y is 0.5βˆ
= 0.5(8.2726) = 4.136 or 4136 visitor days
2
(0.5ˆ ) = (0.5)2 ( ˆ ) = 0.25 ˆ
xx
Var Var
S
\sum 
β β
=
12.43
0.25 = 0.153
20.324
∴ s.d. (0.5βˆ ) = 0.391
∴ standard error of estimate is 391
\item Q16 Comment: Nearly all candidates scored well on parts (i) – (iv), but most used
an inappropriate expression for the standard error of 0.5βˆ .
\end{itemize} 
\newpage

\end{document}
