\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
%%- Quesion 11
A charity sent eight hundred of its supporters information packs about its activities and asked for donations to help it continue its work. Two hundred were sent a pack about its work in education, two hundred a pack about its
work in health care, and four hundred a pack about its anti-poverty programmes. Ninety-seven of the eight hundred people responded with donations; the breakdown is shown in the table below.



\begin{center}
\begin{tabular}{|c|c|c|c|c|}\hline
Expected 	&	Education 	&	Health 	&	Poverty 	&	Total	\\ \hline
Donate & 26 & 31 & 40 & 97 \\ \hline
Don’t donate  & 174 & 169& 360&  703\\ \hline
Total	&	200	&	200	&	400	&	800	\\ \hline
\end{tabular}
\end{center}

Perform a $\chi^2$ test on this table to investigate whether the proportions who
send donations are affected by the type of pack received. 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\medskip 
\begin{itemize}
\item Complete the table of Expected values:
\begin{center}
\begin{tabular}{|c|c|c|c|c|}\hline
Expected 	&	Education 	&	Health 	&	Poverty 	&	Total	\\ \hline
Donate 	&	24.25	&	24.25	&	48.5	&	97	\\ \hline
Don’t donate	&	175.75	&	175.75	&	351.5	&	703	\\ \hline
	&	200	&	200	&	400	&	800	\\ \hline
\end{tabular}
\end{center}

\item Calculate $\chi^2$ = 3.98.
\item The 5\% point of a $\chi^2$ random variable on 2 degrees of freedom is 5.991.
\item We conclude that the $\chi^2$ test is not significant at the 5\% level.
\item 
On the basis of the data collected, it is plausible that the three packs are equally
effective.
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Question 12

A random sample of 200 pairs of observations $(x, y)$ from a discrete bivariate
distribution $(X, Y)$ is as follows:
\begin{itemize}
    \item the observation (-2, 2) occurs 50 times
\item the observation (0, 0) occurs 90 times
\item the observation (2, -1) occurs 60 times.
\end{itemize}

Calculate the sample correlation coefficient for these data.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%- Solution to Question 12
\begin{itemize}
    \item $\sum x = 50(-2) + 0 + 60(2) = 20$
    \item $\sum x2 = 50(4) + 0 + 60(4) = 440$
\item $\sum y = 50(2) + 0 + 60(-1) = 40$ 
\item $\sum y2 = 50(4) + 0 + 60(1) = 260$
\item $\sum xy = 50(-4) + 0 + 60(-2) = -320$
\end{itemize}
so\[ r = \frac{[-320 - (20 \times 40)/200]}{ \sqrt{ [(440 - 202/200)(260 - 402/200)]}} = -0.975\]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{enumerate}
\end{document}
