\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
%% EXAMINATIONS
%% 17 April 2000 (pm)
%% Subject 101 — Statistical Modelling
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\item  An insurance company’s records suggest that experienced drivers (those aged
over 21) submit claims at a rate of 0.1 per year, and inexperienced drivers (those 21 years old or younger) submit claims at a rate of 0.15 per year. 
\begin{itemize}
\item A driver can submit more than one claim a year. The company has 40 experienced and 20 inexperienced drivers insured with it.
\item The number of claims for each driver can be modelled by a Poisson distribution, and claims are independent of each other.
\end{itemize}
 Calculate the
probability the company will receive three or fewer claims in a year. 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\begin{itemize}
\item  As claims are independent the number of claims by inexperienced drivers will
follow a Poisson distribution with mean $20 \times 0.15 = 3$, and the number of claims
made by experienced drivers will follow a Poisson distribution with mean
$40 \times 0.1 = 4$. Again using the independence assumption, the total number of
claims, $X$, is Poisson with mean 7.
\item Thus, 
\begin{eqnarray*}
P(X \leq 3) &=& \sum^{3}_{i=0} e^{-7} \frac{7^i}{i!}\\
& & \\
&=& P(X \leq 0) + P(X \leq 1) + P(X \leq 2) + P(X \leq 3)\\
&=& \left(e^{-7} \frac{7^0}{0!} \right) + \left(e^{-7} \frac{7^1}{1!}\right) + \\
& & \left(e^{-7} \frac{7^2}{2!}\right) + \left(e^{-7} \frac{7^3}{3!}\right)\\
& & \\
&=& 0.082.
\end{eqnarray*}
\item (The answer can also be taken directly from the Green Book, which gives
0.08177.)
\end{itemize}

\newpage

\item The number of claims which arise under a policy of a particular type in a year
is to be modelled as a Poisson(l) random variable. A random sample of 600
such policies gave rise to a total of 72 claims in 1999.
Determine the approximate probability value of this result in a test of
$H0 : l = 0.14$ v $H1 : l < 0.14$ .
\end{enumerate}
\begin{itemize}
\item 6 
Total number of claims $X \sim \mbox{ Poisson}(600\lambda)$

Under H0: $X \sim \mbox{ Poisson}(84) \sim N(84,84)$ approximately

\begin{eqnarray*}
Prob. value &=& P(X \leq 72) \\
&=& P(Z < \frac{(72.5 - 84)}{\sqrt{84}} \\
&=& P(Z < -1.255) \\ 
&=& 0.105\\
\end{eqnarray*}


\end{enumerate}

%%% Question 6


$T \sim \mbox{Poisson}(84)$ is approximated by $T \sim \mbox{N}(84,84)$ 
\begin{eqnarray*}
P(X \leq 73) &=& P(Z \frac{72.5-84}{\sqrt{84}} \mbox{ using continuity correction } \\
&=& P(Z \leq -1.255)\\
&=& \Phi(-1.255) \\
&=& 0.105  \mbox{ from tables }\\
\end{eqnarray*}


\end{itemize}



\end{document}
