
6 (i) Describe a method for generating a sequence U1 U2 of pseudo-random
numbers uniformly distributed in the interval (0 1) and from this a pseudorandom
sequence of numbers uniformly distributed over an arbitrary
range (a b) , where a b . [3]
(ii) Explain the main advantage of using pseudo-random, as opposed to purely
random numbers, for testing the suitability of a model. [1]
(iii) Given a sequence {Un n 1 2 } as in (i), explain how you would simulate
(a) an observation from the Pareto density
1 ( ) = 1 0
(1 )a
a
f x a x
x
(b) a discrete random variable X with probability function given by
1 2
( = ) = for 1 2 ( = 0) =
2 2
P X i i n P X
n n
[4]
(iv) The Pareto distribution in (iii) is often referred to as a heavy-tailed
distribution . Explain the use of this term and discuss under what
circumstances it might offer a suitable model for modelling sizes of claims
arriving at an insurance company. [2]
[Total 10]
103 A2004 5 PLEASE TURN OVER
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
7 (i) Suppose that {Xt : t 0} is a time-homogeneous Markov jump process with
generator matrix A and transition matrix P(t). Let be a (column) vector of
probabilities such that TA = 0T, where 0 is a (column) vector whose
components are all equal to zero.
(a) Prove that
T ( ) = T. d
P t
dt
0
(b) Explain why is known as the stationary distribution for X.
[4]
(ii) A telephone call centre receives calls from customers at an average rate of 0.5
per minute. Each call has a random duration which is exponentially
distributed with mean 3 minutes, independently of the number or duration of
any other calls. Two operators are assigned to handle calls. If a call arrives
when both operators are busy, the call is put on hold unless there are already
two calls on hold, in which case the new call is lost. When a call ends, one of
the calls on hold is immediately put through to the newly free operator.
(a) Identify the five states which are required if this system is to be
modelled as a Markov jump process.
(b) Draw the transition diagram for this system.
(c) Write down the generator matrix for the process.
(d) Evaluate the stationary distribution of the system.
[8]
[Total 12]
103 A2004 6


6 (i) Use a linear congruential generator (LCG). Specify three positive integers
a c m with m a m c and an initial value x0 , then generate a sequence of
x1 x2 xn in the range {0 1 2 m 1} by the recursive rule
Subject 103 (Stochastic Modelling) April 2004 Examiners Report
Page 7
xn (axn 1 c) (modm) n 1 2 N
Then un xn m is a sequence of pseudo-random numbers in [0 1].
To generate a sequence over the range (a b) , a linear transformation
vn (b a)un a is needed. Thus the procedure is: generate a variable
U U(0 1) using a LCG as above; return V (b a)U a .
(ii) The main advantage of using pseudo-random numbers is reproducibility.
When testing a model, this almost certainly depends on a number of
parameters and assumptions, so it is very often desirable to examine the
sensitivity of the model to these assumptions and the values of the parameters
used. It is thus necessary to reduce the random element to a minimum and
rerun the experiment using the same data .
(iii) (a) We use the inverse transform method. The cdf of the Pareto is
0 1
1
( ) 1
(1 ) (1 )
x
a a
a
F x
x x
so the inverse is
F 1( y) (1 y) 1 a 1
Thus the procedure is
Generate an observation U from the Uniform (0,1) distribution.
Return X (1 U) 1 a 1 (or, equally good, X U 1 a 1) as an
observation from the Pareto.
(b) Since the number of possible values for X is n 1, we need to split
the interval (0 1) into n 1 subintervals, the length of which has to be
proportional to the corresponding probabilities. Given a uniform
variable U , one such procedure is:
For i 1 2 n , if (i 1) (n 2) U i (n 2) , then set X i
If U n (n 2) , then set X 0 .
(iv) From (iii), we see that the tail of the Pareto distribution is 1 F(x) (1 x) a ,
which decreases to zero much more slowly than e.g. the exponential or normal
distributions.
Alternatively, it is called fat-tailed because of the relatively high probability
of producing values which are a long way from the mean/median.
Subject 103 (Stochastic Modelling) April 2004 Examiners Report
In general, the Pareto offers a suitable choice for portfolios where there is a
non-negligible chance of very large claims, which is reasonable in various
forms of general insurance, in particular insurance that deals with natural
catastrophes (floods, earthquakes, hurricanes etc).
This question was generally well answered and caused few difficulties.
7 (i) (a) We will use the matrix form of the Kolmogorov DE, which states that
P'(t) AP(t) .
It follows that T ( ) T T ( ) 0T ( ) 0T. d dP
P t AP t P t
dt dt
[The other Kolmogorov DE, P'(t) P(t)A, cannot be used to complete the proof]
(b) If X0 is random with distribution , then the distribution of Xt is given
by TP(t).
The fact that the differential of this is equal to zero implies that TP(t)
= TP(0) = T for all t, in other words Xt has the same distribution for
all t.
For an irreducible Markov chain on a finite state space, the stationary
distribution is also the limiting (equilibrium, long-term, steady state)
distribution.
(ii) (a) The states can be labelled as 0, 1, 2:0, 2:1, 2:2, where 0 and 1 represent
the number of operators occupied and 2:j means that both operators are
occupied and j calls are on hold.
Candidates with a different collection of states can earn 0.5 marks, as
long as their states are states. For example, Call arrives is not a
state, but an event triggering a transition from one state to another.
(b)
The labels on the arcs represent the transition rates.
(c) From the transition diagram, the generator matrix is as below.
0 1 2:0 2:1 2:2
1/2 1/2 1/2 1/2
1/3 2/3 2/3 2/3
Subject 103 (Stochastic Modelling) April 2004 Examiners Report
Page 9
1/ 2 1/ 2 0 0 0
1/ 3 5 / 6 1/ 2 0 0
0 2 / 3 7 / 6 1/ 2 0
0 0 2 / 3 7 / 6 1/ 2
0 0 0 2 / 3 2 / 3
A
(d) We have
0 1
1 1
2 3 1 0
3
2
1 0 2:0
5 1 2
6 2 3 2:0 0
9
8
2:0 1 2:1
7 1 2
6 2 3 2:1 0
27
32
2:1 2:0 2:2
7 1 2
6 2 3 2:2 0
81
128
To solve this we add the condition that the i must sum to 1.
Therefore the stationary distribution is
128 192 144 108 81
653 653 653 653 653
(in decimal form this is
0.1960 0.2940 0.2205 0.1654 0.1240 )
Many candidates made a good attempt at part (i), though not many scored full marks. In part
(ii) a surprising number of candidates were unable to distinguish between states in which
the process stays for a certain length of time and events, which occur instantaneously and
trigger transitions from one state to another.
