
\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
2
A trial for the life of a new battery was carried out by placing 1,000 batteries inside
1,000 fluffy electrical toys. The toys were turned on and left to run for 24 hours. A
researcher returned every hour to count how many toys had stopped operating. On
close analysis it was discovered that some of the toys had ceased to operate due to
mechanical problems and that some of the toys were not switched on in the first place.
Some of the toys were still running after 24 hours.
For each of the following types of censoring, state with reasons whether it is present
in the investigation:
(a)
(b)
(c)
(d)
left-censoring
interval censoring
type I censoring
non-informative censoring

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
2
%%-- Page 4
\begin{itemie}
\item (a) Left-censoring
Not present.
We know for all the toys in the observation exactly when they were turned on.
\item (b) Interval censoring
Present.
We only know the hour during which the toy ceased to function.
\item (c) Type I censoring
Present
Those fluffy toys still running after 24 hours are type I censored.
\item (d) Non-informative censoring
Present
Those not turned on, or failing due to mechanical problems, or censored after
24 hours tell us nothing about battery life.
\end{itemize} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\end{document}
