\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}

%%%%%%%%%%%%%%5

(i) Describe what is meant by homogeneity in a mortality investigation and
explain what problems lack of homogeneity may cause for an insurer in setting
premium rates.

(ii) Suggest six factors in respect of which life office mortality statistics are often
subdivided.

[Total 6]

%%%%%%%%%%%%%5



3
(i)
A population under investigation in a mortality study is homogeneous if all the
lives have similar mortality characteristics.
The problem with lack of homogeneity is that the investigation will reveal
only the average experience of the group.
This means that the insurer will mis-price the policy — for some individuals it
will be too expensive, for others too cheap.
This will be a problem if other insurers do take account of the factor causing
the lack of homogeneity and price correctly. The insurer will be too expensive
for the good risks (and so lose their business) and too cheap for the bad risks
(and so write unprofitable business).
(ii)
Factors usually considered:
·
·
·
·
·
·
·
sex
age
type of policy
smoker/non-smoker status
level of underwriting
duration in force
other factors such as sales channel, policy size, occupation and known
impairments are acceptable
Valid alternative suggestions were credited, but note the word ‘often’ in the
question. Factors which are very rarely used did not receive credit.
