---
title: "Joint Random Variables: Worked Example"
subtitle: "Discrete Probability Distributions"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(tidyquant)
library(ggside)
library(ggplot2)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#0cB2A3",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



\noindent The joint cdf of (X,Y) is
\[
F_{X,Y}(x,y) = 
\begin{cases}
1 – e^{-2x} –e^{-3y} + e^{-(2x+3y)}  &  x,y >0  \\
  0 & \mbox{elsewhere} \\
\end{cases}
\]

\noindent Find
\begin{enumerate}[(a)]
\item  Joint pdf of (X,Y) 
\item Marginal pdf of $X$ and $Y$ 
\item  $P[(X \leq  1) \cap  (Y \leq 1) ] $
\item  $P[(1 < X < 3) \cap  ( 1 < Y < 2)]$
\end{enumerate}


### (a) Joint pdf of (X, Y)

To find the joint pdf, we differentiate the given joint cdf \( F_{X,Y}(x,y) \) with respect to both \(x\) and \(y\):

\[ F_{X,Y}(x,y) = 
\begin{cases}
1 - e^{-2x} - e^{-3y} + e^{-(2x+3y)}, &  x,y > 0 \\
0, & \text{elsewhere}
\end{cases}
\]

The joint pdf \( f_{X,Y}(x,y) \) is obtained by:

\[ f_{X,Y}(x,y) = \frac{\partial^2 F_{X,Y}(x,y)}{\partial x \partial y} \]

\[
\begin{aligned}
f_{X,Y}(x,y) &= \frac{\partial}{\partial y} \left( \frac{\partial F_{X,Y}(x,y)}{\partial x} \right) \\
&= \frac{\partial}{\partial y} \left( 2e^{-2x} - e^{-(2x+3y)} \cdot 3 \right) \\
&= 6 e^{-(2x+3y)}
\end{aligned}
\]

Thus, the joint pdf \( f_{X,Y}(x,y) \) is:

\[
f_{X,Y}(x,y) = 
\begin{cases}
6 e^{-(2x+3y)}, &  x,y > 0 \\
0, & \text{elsewhere}
\end{cases}
\]

### (b) Marginal pdf of \(X\) and \(Y\)

#### Marginal pdf of \(X\):

To obtain the marginal pdf of \(X\), integrate the joint pdf over all \(y\):

\[
f_X(x) = \int_0^\infty f_{X,Y}(x,y) \, dy 
= \int_0^\infty 6 e^{-(2x+3y)} \, dy
\]

Perform the integration:

\[
f_X(x) = 6 e^{-2x} \int_0^\infty e^{-3y} \, dy 
= 6 e^{-2x} \left( \frac{1}{3} \right)
= 2 e^{-2x}
\]

Thus, the marginal pdf of \(X\) is:

\[
f_X(x) = 
\begin{cases}
2 e^{-2x}, &  x > 0 \\
0, & \text{elsewhere}
\end{cases}
\]

#### Marginal pdf of \(Y\):

To obtain the marginal pdf of \(Y\), integrate the joint pdf over all \(x\):

\[
f_Y(y) = \int_0^\infty f_{X,Y}(x,y) \, dx 
= \int_0^\infty 6 e^{-(2x+3y)} \, dx
\]

Perform the integration:

\[
f_Y(y) = 6 e^{-3y} \int_0^\infty e^{-2x} \, dx 
= 6 e^{-3y} \left( \frac{1}{2} \right)
= 3 e^{-3y}
\]

Thus, the marginal pdf of \(Y\) is:

\[
f_Y(y) = 
\begin{cases}
3 e^{-3y}, &  y > 0 \\
0, & \text{elsewhere}
\end{cases}
\]

---

### (c) \( P(X \leq 1 \cap Y \leq 1) \)

To find \( P(X \leq 1 \cap Y \leq 1) \), evaluate the joint cdf at \( x = 1 \) and \( y = 1 \):

\[
P(X \leq 1 \cap Y \leq 1) = F_{X,Y}(1,1) 
= 1 - e^{-2 \cdot 1} - e^{-3 \cdot 1} + e^{-(2 \cdot 1 + 3 \cdot 1)}
= 1 - e^{-2} - e^{-3} + e^{-5}
\]


---

### (d) \( P(1 < X < 3 \cap 1 < Y < 2) \)

To find \( P(1 < X < 3 \cap 1 < Y < 2) \), use the joint cdf:

\[
\begin{aligned}
P(1 < X < 3 \cap 1 < Y < 2) &= F_{X,Y}(3,2) - F_{X,Y}(3,1) - F_{X,Y}(1,2) + F_{X,Y}(1,1) \\
&= \left( 1 - e^{-6} - e^{-6} + e^{-12} \right) - \left( 1 - e^{-6} - e^{-3} + e^{-9} \right) \\
&\quad - \left( 1 - e^{-2} - e^{-6} + e^{-8} \right) + \left( 1 - e^{-2} - e^{-3} + e^{-5} \right) \\
&= e^{-5} - e^{-8} - e^{-9} + e^{-12}
\end{aligned}
\]