---
title: "Moment Generating Functions"
subtitle: "Discrete Probability Distributions"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(tidyquant)
library(ggside)
library(ggplot2)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#0cB2A3",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Let's walk through the steps to find the cumulant generating function, mean, and variance for a given random variable \(X\) with the following probability density function (pdf):

\[ f(x) = \frac{a}{2} e^{-a|x|}, \; a > 0, \; -\infty < x < \infty \]

---
### Step 1: Moment-Generating Function (MGF)

The moment-generating function \(M_X(t)\) is defined as:

\[ M_X(t) = \mathbb{E}[e^{tX}] = \int_{-\infty}^{\infty} e^{tx} f(x) \, dx \]

---

Since \(f(x)\) is different for positive and negative \(x\), we split the integral:

\[ M_X(t) = \int_{-\infty}^{\infty} e^{tx} \frac{a}{2} e^{-a|x|} \, dx \]

\[ = \int_{-\infty}^{0} e^{tx} \frac{a}{2} e^{ax} \, dx + \int_{0}^{\infty} e^{tx} \frac{a}{2} e^{-ax} \, dx \]

For \(x < 0\), \(|x| = -x\):

\[ M_X(t) = \frac{a}{2} \int_{-\infty}^{0} e^{(t+a)x} \, dx + \frac{a}{2} \int_{0}^{\infty} e^{(t-a)x} \, dx \]


---

### Solving the Integrals

For the first integral:

\[ \int_{-\infty}^{0} e^{(t+a)x} \, dx = \left[ \frac{e^{(t+a)x}}{t+a} \right]_{-\infty}^{0} = \frac{1}{t+a} \]

For the second integral:

\[ \int_{0}^{\infty} e^{(t-a)x} \, dx = \left[ \frac{e^{(t-a)x}}{t-a} \right]_{0}^{\infty} = \frac{1}{a-t} \]

Thus,

\[ M_X(t) = \frac{a}{2} \left( \frac{1}{t+a} + \frac{1}{a-t} \right) \]

Simplifying,

\[ M_X(t) = \frac{a}{2} \left( \frac{a-t + t+a}{(t+a)(a-t)} \right) = \frac{a^2}{a^2-t^2} \]

---
### Step 2: Cumulant Generating Function (CGF)

The cumulant generating function \(K_X(t)\) is the natural logarithm of the MGF:

\[ K_X(t) = \log M_X(t) = \log \left( \frac{a^2}{a^2-t^2} \right) = \log(a^2) - \log(a^2-t^2) \]

---

### Step 3: Mean and Variance

The mean and variance are the first and second cumulants, respectively.

#### Mean (\(\mu\)):

The first cumulant is the coefficient of \(t\) in the power series expansion of \(K_X(t)\):

\[ K_X(t) = \log(a^2) - \log(a^2-t^2) = -\log(1 - \frac{t^2}{a^2}) \]

Since there is no linear term in \(t\), the mean \(\mu = 0\).

#### Variance (\(\sigma^2\)):

The second cumulant is the coefficient of \(\frac{t^2}{2}\):

\[ K''_X(t) = \frac{d^2}{dt^2} \left( -\log(1 - \frac{t^2}{a^2}) \right) \]

First derivative:

\[ K'_X(t) = \frac{2t/a^2}{1 - t^2/a^2} \]

Second derivative:

\[ K''_X(t) = \frac{2/a^2 (1 - t^2/a^2) - 2t/a^2 (2t/a^2)}{(1 - t^2/a^2)^2} \]
\[ K''_X(t) = \frac{2/a^2}{(1 - t^2/a^2)^2} \bigg|_{t=0} = \frac{2}{a^2} \]

Thus, the variance \(\sigma^2 = \frac{2}{a^2}\).

---
### Summary

- **Cumulant Generating Function (CGF)**: \( K_X(t) = -\log(1 - \frac{t^2}{a^2}) \)
- **Mean**: \( \mu = 0 \)
- **Variance**: \( \sigma^2 = \frac{2}{a^2} \)


---
