---
title: "Chi Square Goodness of Fit"
subtitle: "Discrete Probability Distributions"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(tidyquant)
library(ggside)
library(ggplot2)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#0cB2A3",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r echo=FALSE,eval=FALSE,include=FALSE}

## ASI - CT3 - 2006 - November - Question 10

```
Suppose that the breakdown of blood types of Blood types in a country follow the following distribution

0 A B AB
50% 35% 10% 5%
In a random sample of 200 stomach cancer patients yielded 
 - 84 having blood type O.
 - 92 having blood type A, 
 - 20 having blood type B, 
 - 4 having blood type AB 


* Are these data significant enough, at 5% level of significance to enable us to reject the
null hypothesis that the blood type distribution of stomach cancer sufferers is the same
as that of the general population?

---

\textbf{Solution}


*  Note: In the question, the distribution of blood groups among the general public is not given. In fact, the
proportion of people with different blood groups is known to be different. 

*  However, in the
absence of this information, the problem has been solved as given below under the assumption of
equality of different proportions. 
*  The law of equal distribution of ignorance is valid if the
proportions of different attributes are not given. 

*  The rejection of such a hypothesis would lead to
different proportions for the various levels of the attributes in the population.
*  $n = 200$
    
    Observed frequencies: $O_A = 92$, $O_B = 20$, $O_{AB} = 4$, $O_O = 84$

*  The expected frequencies: $E_A = 50$, $E_B = 50$, $E_{AB} = 50$, $E_O = 50$ (under $H_0$)

---

*  Calculate the chi-square statistic ($\chi^2$):
    $$
    \chi^2 = \sum \frac{(O_i - E_i)^2}{E_i} = \frac{(92-50)^2}{50} + \frac{(20-50)^2}{50} + \frac{(4-50)^2}{50} + \frac{(84-50)^2}{50}
    $$
    $$
    \chi^2 = \frac{42^2}{50} + \frac{(-30)^2}{50} + \frac{(-46)^2}{50} + \frac{34^2}{50} = 35.28 + 18 + 42.32 + 23.12 = 118.72
    $$

---

*  Critical value of $\chi^2$ with 3 degrees of freedom at 5% level is 7.815
*  Conclusion: Since 118.72 is greater than 7.815, we reject $H_0$. The blood type distribution of people with stomach cancer is significantly different from that of the general public.

