---
title: "Correlation Hypothesis Testing"
subtitle: "Bivariate Analyses"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(tidyquant)
library(ggside)
library(ggplot2)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#0cB2A3",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE,eval=FALSE,include=FALSE}

## ASI - CT3 - 2006 - November - Question 13

```


In a correlation analysis, based on a sample of size 10 from a bivariate normal
distribution, a test of $H_0: \rho = 0$ against $H_1: \rho > 0$ results in a probability value of 0.025. 


**Compute the value of the sample correlation coefficient.**

---

### Solution

Let's break it down:

#### Step 1. **Hypotheses**: 

   - $H_0 : \rho = 0$ 
   - $H_1 : \rho > 0$

#### Step 2. **Test Statistic**: 
   - The test statistic $t$ follows a $t$-distribution with $n-2$ degrees of freedom (df).
   - Test Statistic $$t = \frac{r \sqrt{n-2}}{\sqrt{1-r^2}}$$ with $n = 10$.

#### Step 3. **P-value Interpretation**:
   - Given a one-sided p-value of 0.025, the observed $t$-value is 2.306.
---

#### Step 4. **Solving for $r$**:
   - Plugging in the observed $t$-value, we get:
   $$
   2.306 = \frac{r \sqrt{10-2}}{\sqrt{1-r^2}}
   $$
   - Simplifying:
   $$
   2.306 = \frac{r \sqrt{8}}{\sqrt{1-r^2}}
   $$

---

#### Step 5. **Calculations**:
   - Rearrange and solve for $r$:
   $$
    \frac{2.306}{\sqrt{8}} = \frac{r}{\sqrt{1-r^2}}
   $$
   - Solving
   $$
    \frac{2.306}{\sqrt{8}} = 0.8153
   $$
   
   $$
   0.8153 = \frac{r}{\sqrt{1-r^2}}
   $$
   $$
   r = 0.632
   $$

---

### Fisher's Transformation

Fisher's transformation is a mathematical technique used to transform the sample correlation coefficient $r$ to a variable $Z$ that is approximately normally distributed. 

This transformation makes it easier to conduct hypothesis tests and construct confidence intervals for the correlation coefficient, particularly when the sample size is small.

The transformation is given by:
$$Z = \frac{1}{2} \log \left( \frac{1+r}{1-r} \right)$$
Under the null hypothesis $H_0$, the transformed variable $Z$ follows a normal distribution with mean 0 and standard deviation $\frac{1}{\sqrt{n-3}}$, where $n$ is the sample size. This allows us to perform statistical tests using standard normal distribution properties.

---



### **Using Fisher's Transformation**:
   - Fisher's transformation is given by:
   $$
   Z = \frac{1}{2} \log \left( \frac{1+r}{1-r} \right)
   $$
   - Under $H_0$, $Z$ approximately follows a normal distribution $N(0, \frac{1}{7})$.
   - This also gives $r = 0.630$.

Therefor the sample correlation coefficient $r$ is approximately $0.632$.

---