---
title: "ANOVA"
subtitle: "Analysis of Variance"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
highlightStyle: github
countIncrementalSlides: false
---


```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(tidyquant)
library(ggside)
library(ggplot2)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent_inverse(
  #base_color = "#3C989E")(
  base_color = "#11AA83",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE,eval=FALSE,include=FALSE}

## ASI - CT3 - 2006 - November - Question 12

```


**Problem:**

A private insurance company has three offices at Delhi, Mumbai, and Chennai. To measure how much the employees at these places know about total quality management, a random sample of six employees was selected from each place and given a quality awareness examination. The examination scores are given in the following table.

**Examination Scores:**

| Observation | Delhi | Mumbai | Chennai |
|-------------|-------|--------|---------|
| 1           | 85    | 71     | 59      |
| 2           | 75    | 75     | 64      |
| 3           | 82    | 73     | 62      |
| 4           | 76    | 74     | 69      |
| 5           | 71    | 69     | 75      |
| 6           | 85    | 82     | 67      |


Test the hypothesis that the mean examination score is the same for all three places at the 5% level by stating the model, the hypothesis, and the assumptions involved for this test.

---

### Solution:


*  **Model:**
    
$$x_{ij} = \mu + \alpha_i + e_{ij}$$
where $j = 1,2,\ldots,n_i$; $i = 1,2,3$

- $\mu$: the overall mean
- $e_{ij}$: random component ~ $N(0,\sigma^2)$

---

*  **Hypothesis:**
    
- $H_0 : \alpha_1 = \alpha_2 = \alpha_3$
- $H_1 : \alpha_1 \neq \alpha_2 \neq \alpha_3$

Where $\alpha_i$ (i = 1, 2, 3) are the population means at Delhi, Mumbai, and Chennai, respectively.

*  **Assumptions:**
    
i) Samples are drawn from normal populations  
ii) Samples are independent  
iii) Population variances are equal


---


Let's calculate the overall mean for the three cities (Delhi, Mumbai, and Chennai) based on the examination scores provided.

First, we'll sum up the scores for each city and then calculate the average for each city individually. Finally, we'll compute the overall mean of these three averages.

**Scores:**

| Observation | Delhi | Mumbai | Chennai |
|-------------|-------|--------|---------|
| 1           | 85    | 71     | 59      |
| 2           | 75    | 75     | 64      |
| 3           | 82    | 73     | 62      |
| 4           | 76    | 74     | 69      |
| 5           | 71    | 69     | 75      |
| 6           | 85    | 82     | 67      |

---

**Step-by-Step Calculations:**

1. Sum of the scores for each city:
   - Delhi: $85 + 75 + 82 + 76 + 71 + 85 = 474$
   - Mumbai: $71 + 75 + 73 + 74 + 69 + 82 = 444$
   - Chennai: $59 + 64 + 62 + 69 + 75 + 67 = 396$

2. Number of observations for each city: $6$

3. Mean score for each city:
   - Delhi: $\frac{474}{6} = 79$
   - Mumbai: $\frac{444}{6} = 74$
   - Chennai: $\frac{396}{6} = 66$

4. Overall mean:
   \[ \text{Overall Mean} = \frac{79 + 74 + 66}{3} \approx 73 \]

So, the overall mean for the examination scores from the three cities is approximately $73$.

---

*  **ANOVA:**

$$\begin{array}{|c|c|c|c|c|}
\hline
\text{Sources} & \text{Sum of squares} & \text{d.f.} & \text{MSS} & \text{F} \\ \hline
\text{Treatments} & 516 & 2 & 258 & 9.00 \\ \hline
\text{Error} & 430 & 15 & 28.67 & \\ \hline
\text{Total} & 946 & 17 &  & \\ \hline
\end{array}$$

*  Critical value of $F(2,15)$ at 5% level = 3.682
 
---

*  **Conclusion:**
    
Reject $H_0$. The population means are not equal in these places.

---