\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
4

[5]

(i) Explain the differences between random censoring and Type I censoring in the

context of an investigation into the mortality of life insurance policyholders.

Include in your explanation a statement of the circumstances in which the

censoring will be random, and the circumstances in which it will be Type I. [4]

(ii) Explain what is meant by non-informative censoring in the investigation in (i).

Describe a situation in which censoring might be informative in this

investigation.

%%%%%%%%%%%%%%%%%%%%%%%%

4

(i)

Random censoring occurs when the time at which the ith life is censored is a random variable. The observation will be censored if the censoring time is

less than the (random) lifetime of the life. In an investigation into the mortality of life insurance policyholders, a life will be censored if either the

period of investigation ends before the life dies, or the life withdraws from the

investigation while still alive (perhaps because the policy lapses). Both these mechanisms will generate random censoring.

Type I censoring occurs if the censoring times are known in advance. In most investigations withdrawals do occur, and if it is not known in advance whether

(let alone when) a life will withdraw, then the censoring is not Type I.

If the period of investigation is known in advance and if there are no withdrawals from the investigation while still alive, the censoring will be

Type I.
